---
title: "Outcome Inequalty by Race in Early Stage High Grade Serous Ovarian Cancer"
author: "Kevin Kremer"
date: "`r format(Sys.time(), '%m/%d/%y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(SEER2R)
library(zoo)
library(survival)
library(survminer)
library(kableExtra)
library(RColorBrewer)

#Read SEER case files for ovarian cancer from 2004-2016 into data.frame
ovary.df <- read.SeerStat("ovcase_2004-2016.dic", UseVarLabelsInData = TRUE)

#Change data.frame to tibble
ovary.tib <- as_tibble(ovary.df)

#Rename columns
ovary.tib = rename(ovary.tib,
            Race = "Race_and_origin_recode_NHW_NHB_NHAIAN_NHAPI_Hispanic",
            Histo = "Histologic_Type_ICDO3",
            Lat = "Laterality",
            Summ_Stage = "SEER_Combined_Summary_Stage_2000_2004",
            Stage_Group = "Derived_AJCC_Stage_Group_6th_ed_20042015",
            T_Stage = "Derived_AJCC_T_6th_ed_20042015",
            N_Stage = "Derived_AJCC_N_6th_ed_20042015",
            M_Stage = "Derived_AJCC_M_6th_ed_20042015",
            LND = "Regional_nodes_examined_1988",
            Nodes_Pos = "Regional_nodes_positive_1988",
            Chemo = "Chemotherapy_recode_yes_no/unk",
            DiagYear = "Year_of_diagnosis",
            SurvMonths = "Survival_months",
            COD = "SEER_causespecific_death_classification",
            DiagAge = "Age_at_diagnosis",
            CutoffStatus = "Vital_status_recode_study_cutoff_used",
            DiagMonth = "Month_of_diagnosis_recode",
            CutoffStatus2 = "End_Calc_Vital_Status_Adjusted",
            MonthsFollowed = "Number_of_Intervals_Calculated")

#Remove columns
ovary.tib = select(ovary.tib, -c(Summary_stage_2000_1998, Sex, County, CutoffStatus2, Begin_Calc_Age_Adjusted))

#Change Grades to numeric values
ovary.tib$Grade <- ifelse(ovary.tib$Grade == "Poorly differentiated; Grade III", 3,
        ifelse(ovary.tib$Grade == "Moderately differentiated; Grade II", 2, 
        ifelse(ovary.tib$Grade == "Well differentiated; Grade I", 1,
        ifelse(ovary.tib$Grade == "Undifferentiated; anaplastic; Grade IV", 4, "Unk"))))

#Change Lat to Right/Left/Bil
ovary.tib$Lat <- ifelse(ovary.tib$Lat == "Right - origin of primary", "Right",
        ifelse(ovary.tib$Lat == "Left - origin of primary", "Left", 
        ifelse(ovary.tib$Lat == "Bilateral, single primary", "Bilateral", "Unknown")))

#Change Race to manageable names
ovary.tib$Race <- ifelse(ovary.tib$Race == "Hispanic (All Races)", "Hispanic",
        ifelse(ovary.tib$Race == "Non-Hispanic Asian or Pacific Islander", "API",
        ifelse(ovary.tib$Race == "Non-Hispanic American Indian/Alaska Native", "Native",
        ifelse(ovary.tib$Race == "Non-Hispanic Black", "Black",
        ifelse(ovary.tib$Race == "Non-Hispanic White", "White", "Unk")))))

#Change Nodes_Ex to None, Inadequate, Adequate
ovary.tib$LND <- ifelse(ovary.tib$LND == 0, "None",
        ifelse(ovary.tib$LND > 0 & ovary.tib$LND < 10, "Inadequate", "Adequate"))

#Change Nodes_Pos to Yes/No
ovary.tib$Nodes_Pos <- ifelse(ovary.tib$Nodes_Pos == 0, "No",
                    ifelse(ovary.tib$Nodes_Pos == 98, "Unk", "Yes"))

#Change COD variable to Alive(0) or Dead(1)
ovary.tib$COD <- ifelse(ovary.tib$COD == "Alive or dead of other cause", 0, 1)

#Filter out HGSOC and add column for race and age groups
serous <- filter(ovary.tib, Histo == "8441" | Histo == "8460" | Histo == "8461" )
HGS <- filter(serous, Grade == 3 | Grade == 4)
HGS = mutate(HGS, Black.Race = ifelse(HGS$Race == "Black", "yes", "no") )
HGS = mutate(HGS, Race.Group = ifelse(HGS$Race == "Black", "Black",
                                      ifelse(HGS$Race == "White", "White",
                                             ifelse(HGS$Race == "Hispanic", "Hispanic", "Other"))))
HGS = mutate(HGS, TNM.Stage = ifelse(Nodes_Pos == "No" & M_Stage == "M0", "T1N0M0",
                                     ifelse(Nodes_Pos == "Unk" & M_Stage == "M0", "T1NxM0",
                                            ifelse(Nodes_Pos == "Yes" & M_Stage == "M0", "T1N1M0", "Any M1"))))
HGS = mutate(HGS, Stage_Group_Summ = ifelse(Stage_Group == "IA" | Stage_Group == "IB" | Stage_Group == "IC", "I",
                                            ifelse(Stage_Group == "IIIC", "III",
                                                   ifelse(Stage_Group == "IV", "IV", "Unk"))))
HGS$Age.Group <- cut(HGS$DiagAge, breaks = c(18, 30, 40, 50, 60 ,70, 80, 120), labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80+"), right = FALSE)
HGS.ES <- filter(HGS,(Stage_Group == "IA" | Stage_Group == "IB" | Stage_Group == "IC") & Nodes_Pos != "Yes")
HGS.N0 <- filter(HGS.ES, Nodes_Pos == "No")
HGS.Nx <- filter (HGS.ES, Nodes_Pos == "Unk")
HGS.late <- filter (HGS, Nodes_Pos == "Yes" & M_Stage == "M0")
HGS.mets <- filter(HGS, M_Stage == "M1")
HGS.WB <- filter(HGS.ES, HGS.ES$Race == "Black" | HGS.ES$Race == "White")
HGS.HB <- filter(HGS.ES, HGS.ES$Race == "Black" | HGS.ES$Race == "Hispanic")
HGS.loc <- filter(HGS, Summ_Stage == "Localized only" & Nodes_Pos != "Yes")
HGS.SG <- filter(HGS, Stage_Group_Summ != "Unk")
```

# Survival of early-stage HGSOC by race

```{r, echo = FALSE}
HGS.ES$Race <- factor(HGS.ES$Race, levels = c('White', 'Black', 'Hispanic', 'API', 'Native'))
fit.byRace <- survfit(Surv(SurvMonths, COD) ~ Race, HGS.ES)
ggsurvplot(fit.byRace, HGS.ES, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Race", title = "Survival of early stage HGSOC patients stratified by Race", legend.labs = c("White", "Black", "Hispanic", "Asian/Pacific Islander", "Native American"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
pairwise_survdiff(Surv(SurvMonths, COD) ~ Race, HGS.ES, rho = 0)
HGS.ES.Racetbl <- table(HGS.ES$Race)
kbl(HGS.ES.Racetbl, col.names = c("Race", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

# Comparing Black Race to All Other Races Combined

```{r, echo = FALSE}
fit.byBlack.Race <- survfit(Surv(SurvMonths, COD) ~ Black.Race, HGS.ES)
ggsurvplot(fit.byBlack.Race, HGS.ES, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Race", title = "Survival of early stage HGSOC patients stratified by Race", legend.labs = c("Other", "Black"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.ES.Black.Racetbl <- table(HGS.ES$Black.Race)
kbl(HGS.ES.Black.Racetbl, col.names = c("Black Race", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.Black.Race <- coxph(Surv(SurvMonths, COD) ~ Black.Race, HGS.ES)
ggforest(ph.Black.Race, HGS.ES)
summary(ph.Black.Race)
```

# Comparing Black Race to White Race

```{r, echo = FALSE}
HGS.WB$Race <- factor(HGS.WB$Race, levels = c('White', 'Black'))
fit.byWB <- survfit(Surv(SurvMonths, COD) ~ Race, HGS.WB)
ggsurvplot(fit.byWB, HGS.WB, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Race", title = "Survival of early stage HGSOC patients stratified by Race", legend.labs = c("White", "Black"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.ES.WBtbl <- table(HGS.WB$Race)
kbl(HGS.ES.WBtbl, col.names = c("Race", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.WB <- coxph(Surv(SurvMonths, COD) ~ Race, HGS.WB)
ggforest(ph.WB, HGS.WB)
summary(ph.WB)
```

# Comparing Black Race to Hispanic Race

```{r, echo = FALSE}
HGS.HB$Race <- factor(HGS.HB$Race, levels = c('Hispanic', 'Black'))
fit.byHB <- survfit(Surv(SurvMonths, COD) ~ Race, HGS.HB)
ggsurvplot(fit.byHB, HGS.HB, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Race", title = "Survival of early stage HGSOC patients stratified by Race", legend.labs = c("Hispanic", "Black"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.ES.HBtbl <- table(HGS.HB$Race)
kbl(HGS.ES.HBtbl, col.names = c("Race", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.HB <- coxph(Surv(SurvMonths, COD) ~ Race, HGS.HB)
ggforest(ph.HB, HGS.HB)
summary(ph.HB)
```

# Does the addition of chemotherapy in patients with unknown nodal status improve outcomes in different races?

## Black Race

```{r, echo = FALSE}
HGS.ES.Black.Chemo = filter(HGS.ES, Race == "Black" & Chemo == "Yes")
fit.Black.Chemo = survfit(Surv(SurvMonths, COD) ~ Nodes_Pos, HGS.ES.Black.Chemo)
ggsurvplot(fit.Black.Chemo, HGS.ES.Black.Chemo, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Stage", title = "Survival of Black early stage HGSOC patients that \nreceived chemotherapy stratified by Stage", legend.labs = c("T1N0M0", "T1NxM0"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.ES.Black.Chemotbl <- table(HGS.ES.Black.Chemo$Nodes_Pos)
kbl(HGS.ES.Black.Chemotbl, col.names = c("Positive Nodes", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.Black.Chemo <- coxph(Surv(SurvMonths, COD) ~ Nodes_Pos, HGS.ES.Black.Chemo)
ggforest(ph.Black.Chemo, HGS.ES.Black.Chemo)
summary(ph.Black.Chemo)
```

## White Race

```{r, echo = FALSE}
HGS.ES.White.Chemo = filter(HGS.ES, Race == "White" & Chemo == "Yes")
fit.White.Chemo = survfit(Surv(SurvMonths, COD) ~ Nodes_Pos, HGS.ES.White.Chemo)
ggsurvplot(fit.White.Chemo, HGS.ES.White.Chemo, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Stage", title = "Survival of White early stage HGSOC patients that \nreceived chemotherapy stratified by Stage", legend.labs = c("T1N0M0", "T1NxM0"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.ES.White.Chemotbl <- table(HGS.ES.White.Chemo$Nodes_Pos)
kbl(HGS.ES.White.Chemotbl, col.names = c("Positive Nodes", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.White.Chemo <- coxph(Surv(SurvMonths, COD) ~ Nodes_Pos, HGS.ES.White.Chemo)
ggforest(ph.White.Chemo, HGS.ES.White.Chemo)
summary(ph.White.Chemo)
```

## Hispanic Race

```{r, echo = FALSE}
HGS.ES.Hisp.Chemo = filter(HGS.ES, Race == "Hispanic" & Chemo == "Yes")
fit.Hisp.Chemo = survfit(Surv(SurvMonths, COD) ~ Nodes_Pos, HGS.ES.Hisp.Chemo)
ggsurvplot(fit.Hisp.Chemo, HGS.ES.Hisp.Chemo, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Stage", title = "Survival of Hispanic early stage HGSOC patients that \nreceived chemotherapy stratified by Stage", legend.labs = c("T1N0M0", "T1NxM0"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.ES.Hisp.Chemotbl <- table(HGS.ES.Hisp.Chemo$Nodes_Pos)
kbl(HGS.ES.Hisp.Chemotbl, col.names = c("Positive Nodes", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.Hisp.Chemo <- coxph(Surv(SurvMonths, COD) ~ Nodes_Pos, HGS.ES.Hisp.Chemo)
ggforest(ph.Hisp.Chemo, HGS.ES.Hisp.Chemo)
summary(ph.Hisp.Chemo)
```

# Does use of chemotherapy matter by stage for each race?

## Black Race

```{r, echo = FALSE}
HGS.Black.N0 <- filter(HGS.ES, HGS.ES$Race == "Black" & HGS.ES$Nodes_Pos == "No")
fit.Black.N0.byChemo = survfit(Surv(SurvMonths, COD) ~ Chemo, HGS.Black.N0)
ggsurvplot(fit.Black.N0.byChemo, HGS.Black.N0, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Chemotherapy received", title = "Survival of Black T1N0M0 patients stratified \nby receipt of chemotherapy", legend.labs = c("No/Unknown", "Yes"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.Black.N0.Chemotbl <- table(HGS.Black.N0$Chemo)
kbl(HGS.Black.N0.Chemotbl, col.names = c("Chemotherapy received", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.Black.N0 <- coxph(Surv(SurvMonths, COD) ~ Chemo, HGS.Black.N0)
ggforest(ph.Black.N0, HGS.Black.N0)
summary(ph.Black.N0)
```


```{r, echo = FALSE}
HGS.Black.Nx <- filter(HGS.ES, HGS.ES$Race == "Black" & HGS.ES$Nodes_Pos == "Unk")
fit.Black.Nx.byChemo = survfit(Surv(SurvMonths, COD) ~ Chemo, HGS.Black.Nx)
ggsurvplot(fit.Black.Nx.byChemo, HGS.Black.Nx, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Chemotherapy received", title = "Survival of Black T1NxM0 patients stratified \nby receipt of chemotherapy", legend.labs = c("No/Unknown", "Yes"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.Black.Nx.Chemotbl <- table(HGS.Black.Nx$Chemo)
kbl(HGS.Black.Nx.Chemotbl, col.names = c("Chemotherapy received", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.Black.Nx <- coxph(Surv(SurvMonths, COD) ~ Chemo, HGS.Black.Nx)
ggforest(ph.Black.Nx, HGS.Black.Nx)
summary(ph.Black.Nx)
```

## White Race

```{r, echo = FALSE}
HGS.White.N0 <- filter(HGS.ES, HGS.ES$Race == "White" & HGS.ES$Nodes_Pos == "No")
fit.White.N0.byChemo = survfit(Surv(SurvMonths, COD) ~ Chemo, HGS.White.N0)
ggsurvplot(fit.White.N0.byChemo, HGS.White.N0, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Chemotherapy received", title = "Survival of White T1N0M0 patients stratified \nby receipt of chemotherapy", legend.labs = c("No/Unknown", "Yes"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.White.N0.Chemotbl <- table(HGS.White.N0$Chemo)
kbl(HGS.White.N0.Chemotbl, col.names = c("Chemotherapy received", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.White.N0 <- coxph(Surv(SurvMonths, COD) ~ Chemo, HGS.White.N0)
ggforest(ph.White.N0, HGS.White.N0)
summary(ph.White.N0)
```


```{r, echo = FALSE}
HGS.White.Nx <- filter(HGS.ES, HGS.ES$Race == "White" & HGS.ES$Nodes_Pos == "Unk")
fit.White.Nx.byChemo = survfit(Surv(SurvMonths, COD) ~ Chemo, HGS.White.Nx)
ggsurvplot(fit.White.Nx.byChemo, HGS.White.Nx, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Chemotherapy received", title = "Survival of White T1NxM0 patients stratified \nby receipt of chemotherapy", legend.labs = c("No/Unknown", "Yes"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.White.Nx.Chemotbl <- table(HGS.White.Nx$Chemo)
kbl(HGS.White.Nx.Chemotbl, col.names = c("Chemotherapy received", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.White.Nx <- coxph(Surv(SurvMonths, COD) ~ Chemo, HGS.White.Nx)
ggforest(ph.White.Nx, HGS.White.Nx)
summary(ph.White.Nx)
```

## Hispanic

```{r, echo = FALSE}
HGS.Hisp.N0 <- filter(HGS.ES, HGS.ES$Race == "Hispanic" & HGS.ES$Nodes_Pos == "No")
fit.Hisp.N0.byChemo = survfit(Surv(SurvMonths, COD) ~ Chemo, HGS.Hisp.N0)
ggsurvplot(fit.Hisp.N0.byChemo, HGS.Hisp.N0, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Chemotherapy received", title = "Survival of Hispanic T1N0M0 patients stratified \nby receipt of chemotherapy", legend.labs = c("No/Unknown", "Yes"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.Hisp.N0.Chemotbl <- table(HGS.Hisp.N0$Chemo)
kbl(HGS.Hisp.N0.Chemotbl, col.names = c("Chemotherapy received", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.Hisp.N0 <- coxph(Surv(SurvMonths, COD) ~ Chemo, HGS.Hisp.N0)
ggforest(ph.Hisp.N0, HGS.Hisp.N0)
summary(ph.Hisp.N0)
```


```{r, echo = FALSE}
HGS.Hisp.Nx <- filter(HGS.ES, HGS.ES$Race == "Hispanic" & HGS.ES$Nodes_Pos == "Unk")
fit.Hisp.Nx.byChemo = survfit(Surv(SurvMonths, COD) ~ Chemo, HGS.Hisp.Nx)
ggsurvplot(fit.Hisp.Nx.byChemo, HGS.Hisp.Nx, pval = TRUE, xlab = "Months", break.time.by = 12, break.y.by = 0.1, legend.title = "Chemotherapy received", title = "Survival of Hispanic T1NxM0 patients stratified \nby receipt of chemotherapy", legend.labs = c("No/Unknown", "Yes"), legend = "bottom", palette = "Set1", ggtheme = theme_gray())
HGS.Hisp.Nx.Chemotbl <- table(HGS.Hisp.Nx$Chemo)
kbl(HGS.Hisp.Nx.Chemotbl, col.names = c("Chemotherapy received", "Count")) %>% 
        kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
```{r, echo = FALSE}
ph.Hisp.Nx <- coxph(Surv(SurvMonths, COD) ~ Chemo, HGS.Hisp.Nx)
ggforest(ph.Hisp.Nx, HGS.Hisp.Nx)
summary(ph.Hisp.Nx)
```

# Overall CoxPH and Forest plot

```{r, echo = FALSE}
HGS.ES$Race.Group <- factor(HGS.ES$Race.Group, levels = c('White', 'Hispanic', 'Black', 'Other'))
ph.ES <- coxph(Surv(SurvMonths, COD) ~ TNM.Stage + Lat + Chemo + Race.Group + LND, data = HGS.ES)
ggforest(ph.ES, HGS.ES)
summary(ph.ES)
```